This is the title [Article v\versionnumber]

### 2022-04-26 manuscript/manuscript.tex

\subsubsection{Sixth Order Entropy}
\begin{equation}
    S_\textit{k} \approx \frac{-R}{N_\textit{k}} \left( \sum _{i=1}^{N_k} ln g_{NN, \textit{i}} \left( \textbf{r}, \mathbf {\omega} \right) + \gamma \right)
\end{equation}

\subsection{Solvation Energy}
\subsubsection{PME implementation}

@@

After choosing an appropriate reference value, we need to subtract this value from Eww.
Make sure to subtract the reference from the normalized (\_norm) data.
In \software{gisttools}, this is done simply by setting \inlinecode{gist.eww\_ref}.
In \software{gistpp}, this can be done as follows: \todo{This might become invalid with future GIST versions, depending on whether we adopt the COM as molecular center.} \todo{Better use the population.}

@@

It might also be interesting to visualize the average energy of a water molecule at each grid voxel.
This quantity is given by 2*Eww + Esw.
The energy referencing leads to non-zero normalized values where the population is zero.
While this is irrelevant for further post-processing, it is advantageous for visualization to set those empty regions to zero.
A .dx file can be produced using gisttools as follows:

\begin{lstlisting}[style=python]
import gisttools as gt
gist = gt.gist.load_gist_file("gist.dat",
    struct="solute-centered.pdb", eww_ref=-9.533)
gist['E_norm'] = gist['Eww_norm'] * 2 + gist['Esw_norm']
gist.loc[gist['population'] == 0, 'E_norm'] = 0
gist.save_dx('E_norm', 'gist-E-per-mol-norm.dx')
\end{lstlisting}

Visualize the free energy, the total energy, and the entropy, at several iso-levels.
An minimal PyMOL input might look like this:

\begin{lstlisting}
load solute-centered.pdb
show sticks, resname BTN
as surface, not resname BTN
load gist-gO.dx, gO
isosurface gO_high, gO, 2.
\end{lstlisting}

Adapt the above PyMOL script to visualize the average energy per solvent at a positive and a negative isolevel, as well as regions of negative entropy per solvent.
The expected representation of the streptavidin binding pocket is shown in Figure \ref{fig_binding_pocket_pymol}.

\begin{figure}
	\centering
	\includegraphics[width=0.8\linewidth]{figures/binding_pocket_S_-2_E_-2_E_3.png}
	\caption{Free energy contributions in the streptavidin binding pocket. Red: high solvent energy (at +3 kcal/mol). Blue: low solvent energy (at -2 kcal/mol). Green: low solvent entropy $T\Delta S^{six}$ (at -2 kcal/mol).}
	\label{fig_binding_pocket_pymol}
\end{figure}

You should find that there are several regions in the binding pocket that feature a high solvent energy and/or low solvent entropy.
However, such regions also exist outside of the binding pocket, and there are also low-energy regions within the binding pocket.
Therefore, the strong affinity to biotin can not be explained solely by the hydration contribution, and the direct interaction energy also needs to be taken into account.

\subsection{Contribution of Hydration to Binding}
In the next step, calculate the $\dgsolv$ contributions around the biotin ligand in each of the three systems (biotin, streptavidin, and complex).
Note that our integration region does not fully reach into bulk, but also comprises regions that are close to the streptavidin protein. 
Therefore, our results for each system will depend on the exact position of the integration boundary. 
To avoid inconsistencies, it is important to choose exactly the same integration region for each system.
Although we kept the systems rigid, the center of mass might have been shifted during pressure equilibration.
Therefore, we align the complex structure to the streptavidin system and use the shifted biotin coordinates to define the integration region.

We recommend using a \software{Jupyter Notebook} for the following analyses, although they can also be done in a script or an interactive Python prompt.
The code is not explained in detail here, but should be clear after reading the previous sections.

Load the files and double check the frame numbers and reference density.
\begin{lstlisting}[style=python]
import numpy as np
from gisttools.gist import load_gist_file
import matplotlib.pyplot as plt

compl = load_gist_file('complex/gist/gist.dat', \
    struct='complex/gist/solute-centered.pdb')
print(compl.n_frames, compl.rho0)
biotin = load_gist_file('biotin/gist/gist.dat', \
    struct='biotin/gist/solute-centered.pdb')
print(biotin.n_frames, biotin.rho0)
strept = load_gist_file('streptavidin/gist/gist.dat', \
    struct='streptavidin/gist/solute-centered.pdb')
print(strept.n_frames, strept.rho0)
\end{lstlisting}

Assign reference energies and check them for plausibility.
\begin{lstlisting}[style=python]
biotin.eww_ref = biotin.detect_reference_value()
print("Biotin:", biotin.eww_ref)
strept.eww_ref = strept.detect_reference_value()
print("Streptavidin:", strept.eww_ref)
compl.eww_ref = compl.detect_reference_value()
print("Complex:", compl.eww_ref)
\end{lstlisting}

Subtract a reference entropy from the dTSsix columns.
\begin{lstlisting}[style=python]
def reference_entropy(gf):
    if 'dTSsix_unref_norm' not in gf.data.columns:
        gf['dTSsix_unref_norm'] = gf['dTSsix_norm']
        gf['dTSsix_unref_dens'] = gf['dTSsix_dens']
    gf['dTSsix_norm'] = gf['dTSsix_unref_norm'] \
        - gf.detect_reference_value('dTSsix_unref_dens')
    gf['dTSsix_dens'] = gf.norm2dens(gf['dTSsix_norm'])
reference_entropy(biotin)
reference_entropy(strept)
reference_entropy(compl)
\end{lstlisting}

Compute the atom positions that define the integration region.
For the streptavidin integral, we align the complex structure onto the streptavidin structure and then use the biotin positions as centers.
Note that we use \software{mdtraj} here, since \software{gisttools} stores the reference structure as a \inlinecode{mdtraj.Trajectory} object, but any other library for post-processing of trajectories works as well.
\begin{lstlisting}[style=python]
col = 'dTSsix_dens'
def select(traj, sel):
    "Slice a Trajectory by selection mask."
    return traj.atom_slice(traj.top.select(sel))
# we multiply by 10 to convert nm to Angstrom.
compl_x = select(compl.struct, biotin_mask).xyz[0] * 10.
biotin_x = select(biotin.struct, biotin_mask).xyz[0] * 10.
aligned = compl.struct[:].superpose(strept.struct, \
    atom_indices=strept.struct.top.select(strept_mask))
aligned = select(aligned, biotin_mask)
strept_x = aligned.xyz[0] * 10.

bins, biotin_rdf = biotin.rdf( \
    col, centers=biotin_x, bins=100, rmax=24)
bins, strept_rdf = strept.rdf( \
    col, centers=strept_x, bins=100, rmax=24)
bins, compl_rdf = compl.rdf( \
    col, centers=compl_x, bins=100, rmax=24)
\end{lstlisting}

Now, subtract the monomer rdfs from the complex, and compute the integrals.
If you also visualize the individual rdfs, you will notice that the difference converges much better with increasing radius than the contributions.
This is because the cutoff contains regions that are close to the atom, since the integration region does not comprise the whole protein.
\begin{lstlisting}[style=python]
difference = compl_rdf - biotin_rdf - strept_rdf
cutoff = 12
integral = difference[bins < cutoff].sum()
print("Integral = {}".format(integral))
plt.plot(bins, np.cumsum(difference))
plt.axvline(cutoff)
plt.xlabel('distance to biotin [A]')
plt.ylabel('dG contribution [kcal/mol]')
\end{lstlisting}
%\section{Content and links}
%
%A tutorial will normally draw on additional files and materials; clearly indicate where and how these are available, with links, and how they are being archived for the long-term and maintained so they stay current.
%You will likely want to reference your GitHub repository as a central point to access all of this information, and then the GitHub repository may link out to other content as needed.

@@

\section*{Potentially Conflicting Interests}
%%%%%%%j
%Declare any potentially competing interests, financial or otherwise
%%%%%%%

>>>>

The second method directly calculates the solvation entropy by evaluating the six-dimensional integral (3 for position and 3 for orientation) 
using the nearest-neighbor approach. Unlike the first method, the nearest neighbor estimate for the sixth-order entropy considers all the water 
found within the same and neighboring voxels.  
\begin{equation}
	S_\textit{k} \approx \frac{-R}{N_\textit{k}} \left( \sum _{i=1}^{N_k} ln g_{NN, \textit{i}} \left( \textbf{r}, \mathbf {\omega} \right) + \gamma \right)
\end{equation}

\subsection{Solvation Energy}
The solvation energy is readily calculated by summing up the contributions of the water-water interaction and the water-solute interaction.
\begin{equation}
	\Delta E_\textit{solv} = \Delta E_\textit{sw} + \Delta E_\textit{ww}
\end{equation}

\begin{equation}
	\Delta E_\textit{sw} \equiv \frac{1}{8\pi^2} \int g_\textit{sw}\left(\omega|\textbf{r}\right) U_\textit{sw}\left(\textbf{r}, \omega\right) d\omega
\end{equation}

Similar to the entropy integrals, the solute-water solvation energy integral also decays with increasing distance from the solute. Hence, the solvation energy can be approximated by local spatial integrals. However, there is a complication arising from this approximation. Water molecules outside the region of interest will also be accounted for in the pairwise water term. For example, given two different regions $R_a$ and $R_b$, will not sum up to the total energy of the two regions due to double counting.

\begin{equation}
\begin{aligned}
	\Delta E_\textit{ww} \equiv & \left(\frac{1}{8\pi^2} \right)^2 \rho^o \int g_\textit{sw}\left(\omega|\textbf{r}\right) \\
	& \times \left[g_\textit{sw}\left(\textbf{r}^\prime , \omega^\prime \right) - g^o_\textit{ww} \left(\textbf{r}, \omega, \textbf{r}^\prime, \omega^\prime \right)\right] \\
	& \times U_\textit{ww}\left(\textbf{r}, \omega, \textbf{r}^\prime, \omega^\prime\right)d\omega d\textbf{r}^\prime d\omega^\prime
\end{aligned}
\end{equation}
There are two ways in which this double counting was accounted for in the original paper. First, using the water-displacement method. This method is applicable when a determining the energetic cost of kicking out a water molecule. 
\begin{equation}
	\Delta E^R_\textit{ww} = n^\textit{R} E^\textit{bulk}_\textit{ww} - E^\textit{R,corr}_\textit{ww}
\end{equation}

The second method uses normalized-water properties to account the difference in water properties of two different regions. The normalized water property is defined as the water property per water molecule. 
\begin{equation}
	\Delta E^\textit{R,norm}_\textit{ww} = E^\textit{R,norm}_\textit{ww} - 2E^\textit{bulk}_\textit{ww}
\end{equation}

\subsection{PME implementation}

@@

After choosing an appropriate reference value, we need to subtract this value from Eww.
Make sure to subtract the reference from the normalized (\_norm) data.
In \software{gisttools}, this is done simply by setting \inlinecode{gist.eww\_ref}.
In \software{gistpp}, this can be done as follows: \todo{This might become invalid with future GIST versions, depending on whether we adopt the COM as molecular center.}

@@

Visualize the free energy, the total energy, and the entropy, at several iso-levels.
The expected result is shown in Figure ???.

\subsection{Contribution of Hydration to Binding}

%\section{Content and links}
%
%A tutorial will normally draw on additional files and materials; clearly indicate where and how these are available, with links, and how they are being archived for the long-term and maintained so they stay current.
%You will likely want to reference your GitHub repository as a central point to access all of this information, and then the GitHub repository may link out to other content as needed.

@@

\section*{Potentially Conflicting Interests}
%%%%%%%
%Declare any potentially competing interests, financial or otherwise
%%%%%%%
